Bacon, Eggs, and Inflation: Can household groceries track inflation?

### 2023-04-21 outputs/paper/paper.qmd
Paper updates
- Fixed typos
- Rendered without model section

### Pasta Sauce (per 650 millilitres)
As a shelf-stable food item, pasta sauce possessed the opportunity to display an ability to track or resist inflation. Despite these assumptions, pasta sauce generally increased at rates greater than that of yearly inflation. On average, the cost of eggs increased by 5.71% over the observed period while the inflation rate increased by 3.04%. The highest price increase was between 2021 and 2022, which saw an increase 12.7% from \$2.44 to \$2.75. Generally, aside from 2018 when the cost increase was below interest rates, the cost of pasta sauce remained above the rate of inflation. In 2020 and in 2022, the difference was 6%.

@@

In the given time period, generally pasta sauce remained above yearly inflation rates and is not suitable for tracking Canada's inflation rate. With the cost increase peaking at 6% in two of the five observed years, pasta sauce defied expectations by consistently remaining at rates above inflation.

@@

One example of an external factor is "food inflation," or the impact of climate change and high transportation costs, "as a result of higher oil prices" [@cite_food_price_report]. When consider the raw cost of food, examining the entire supply chain is required to fully understand the cost. Changes in production, transport, and sellers may increase---or decrease---the cost of food beyond the influence of inflation.

Another example of an external factor is "greedflation," or the "taking advantage of high inflationary times to earn excessive profits at the expense of consumers who have limited choice in grocery shopping" [@cite_food_price_report]. Large Canadian grocery chains are currently in the spotlight for price gouging, taking advantage of the high costs created by the recent pandemic. This question is so significant that Canada's Competition Bureau launched a study of grocery store competition in October 2022, examining "to what extent are higher grocery prices a result of changing competitive dynamics" [@cite_competition_bureau]. Roughly 80% of Canadians claim "abuse in the system" by grocery stores, placing the blame for high costs squarely on corporations [@cite_toronto_sun]. A review of the largest Canadian grocery chains found "flat" performance in the income of these chains, "when compared to the increase in the cost of living," and "yields were usually below or at the same level as the rate of food inflation" in the observed 5-year period [@cite_toronto_sun].

>>>>

### Pasta Sauce (per 650 millilitres)
As a shelf-stable food item, pasta sauce possessed the opportunity to track or resist inflation. Despite these assumptions, pasta sauce generally increased at rates greater than yearly inflation. On average, the cost of eggs increased by 5.71% over the observed period while the inflation rate increased by 3.04%. The highest price increase was between 2021 and 2022, which saw an increase 12.7% from \$2.44 to \$2.75. Generally, aside from 2018 when the cost increase was below interest rates, the cost of pasta sauce increased above the rate of inflation. In 2020 and in 2022, the difference was 6%.

@@

In the given time period, pasta sauce generally remained above yearly inflation rates and is not suitable for tracking Canada's inflation rate. With the cost increase peaking at 6% in two of the five observed years, pasta sauce defied expectations by consistently remaining at rates above inflation.

@@

One example of an external factor is "food inflation," or the impact of climate change and high transportation costs, "as a result of higher oil prices" [@cite_food_price_report]. When considering the raw cost of food, examining the entire supply chain is required to fully understand the cost. Changes in production, transport, and sellers may increase---or decrease---the cost of food beyond the influence of inflation.

Another example of an external factor is "greedflation," or "taking advantage of high inflationary times to earn excessive profits at the expense of consumers who have limited choice in grocery shopping" [@cite_food_price_report]. Large Canadian grocery chains are currently in the spotlight for price gouging, taking advantage of the high costs created by the recent pandemic. This question is so significant that Canada's Competition Bureau launched a study of grocery store competition in October 2022, examining "to what extent are higher grocery prices a result of changing competitive dynamics" [@cite_competition_bureau]. Roughly 80% of Canadians claim "abuse in the system" by grocery stores, placing the blame for high costs squarely on corporations [@cite_toronto_sun]. A review of the largest Canadian grocery chains found "flat" performance in the income of these chains, "when compared to the increase in the cost of living," and "yields were usually below or at the same level as the rate of food inflation" in the observed 5-year period [@cite_toronto_sun].
